### 2015-12-28 cover_letter.tex
adding to cover letter

ANGSD-wrapper is intended to simplify analysis and visualization of next generation resequencing data.

Visualization of analysis results is implemented using Shiny graphics in R.

The majority of the code is written in Bash, with graphics implemented in R. For speed, one sorting routine is implemented in Python.