### 2016-01-29 cover_letter.tex
cl updates

Dear Managing Editor,

Enclosed you will find our manuscript, "ANGSD-wrapper: utilities for analyzing next generation sequencing data."
This manuscript reports the novel software, ANGSD-wrapper, which is intended to simplify analysis and visualization of next generation resequencing data.
The focus of this program is on usability and it extends the previously published software, ANGSD and programs related to it. 
ANGSD-wrapper consists of two parts: 1) ANGSD pipelines that simplify multi-step analyses into a single step, and 2) interactive graphing based on the direct output of ANGSD-wrapper. 
Interactive visualization of analysis results is implemented using Shiny graphics in R, which uses modern web technologies such as Javascript and HTML5. 
The majority of the code is written in Bash, with graphics implemented in R. For speed, one sorting routine is implemented in Python.
We demonstrate the usefulness of the software by analyzing previously published data in maize and teosinte.

We believe that usability and usefulness of our software will appeal to the readership of Molecular Ecology Resources.

Thank you,

Arun Durvasula and Jeffrey Ross-Ibarra