### 2016-01-30 cover_letter.tex
updating cover letter

Dear Managing Editor,

Enclosed you will find our manuscript, "ANGSD-wrapper: utilities for analyzing next generation sequencing data." This manuscript reports the novel software, "ANGSD-wrapper," which simplifies analysis and visualization of next generation resequencing data. The focus of this program is on usability and it extends utlity of the previously published software, "ANGSD" and programs related to it.

ANGSD-wrapper consists of two parts: 1) ANGSD pipelines that simplify multi-step analyses into a single step, and 2) interactive graphing based on the output of ANGSD-wrapper. Interactive visualization of results is implemented using Shiny graphics in R, which uses modern web technologies such as Javascript and HTML5.

The majority of the code is written in Bash, with graphics implemented in R. For speed, one sorting routine is implemented in Python.
We demonstrate the usefulness of the software by analyzing previously published data in maize and teosinte.

The usability and usefulness of our software should appeal to the readership of Molecular Ecology Resources.

Suggested reviewers for the manuscript include:
	- Sung Chun; Brigham and Women’s Hospital and Harvard Medical School; sgchun@partners.org 
	- Naoki Takebayashi; University of Alaska, Fairbanks; ntakebayashi@alaska.edu

Thank you,

Arun Durvasula and Jeffrey Ross-Ibarra